---
title: "EAS 648 Lab01"
author: "Yinjiao Zhong"
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "/Users/zhongyinjiao/Desktop/EAS648")
```


```{r cars}
library(ggplot2)
library(sf)
library(tidyverse)
```

## LabLecture

```{r}
boulder <- st_read("/Users/zhongyinjiao/Desktop/EAS648/Lab01/BoulderSocialMedia.shp")
boulder
```
```{r}
ggplot() +
    geom_sf(data =boulder,
    fill = NA, alpha = .2) +
    theme_bw()
```
```{r}
boulder = st_transform(boulder, 26753) 
ggplot() +
    geom_sf(data =boulder,
    fill = NA, alpha = .2) +
    theme_bw()
```
```{r}
ggplot() +
    geom_sf(data =boulder, aes(color=PT_Elev),
    fill = NA, alpha = .2) +
    theme_bw()
```
```{r}
ggplot() +
    geom_sf(data =boulder, aes(color=PT_Elev),
    fill = NA, alpha = .2) +
  scale_colour_gradientn(colours = terrain.colors(10)) +  
  theme_bw()
```
```{r}
boulder %>%
    mutate(high_elev = ifelse(PT_Elev >= 2200, TRUE, FALSE))%>% 
ggplot() +
  geom_sf(aes(color=high_elev),
    fill = NA, alpha = .2)  +  
  theme_bw()
```
```{r}
boulder %>%
  filter(DB ==  'Pano' | DB == 'Flickr') %>%
  ggplot(aes(x=DB, y=Street_dis)) + 
  geom_boxplot()
```
```{r}
library(sf)
library(ggspatial)
library(viridis)
## the function gives the hexadecimal colors 
## the interger give the numbers of colors
magma(10)
```
```{r}
ggplot() +
    geom_sf(data = boulder, aes(color=PT_Elev),
    fill = NA, alpha = .2) + 
    scale_colour_gradientn(colours = magma(10))
```
```{r}
summary(boulder$DB)
```
```{r}
p <- ggplot() +
  annotation_spatial(boulder) +
  layer_spatial(boulder, aes(col = DB))
p + scale_color_brewer(palette = "Dark2")
```
```{r}
library(tmap)
tmap_mode("plot")  
## Add the data - these are specific to the vector or raster
  ## which variable, is there a class interval, palette, and other options
tm_shape(boulder) + 
  tm_symbols(col = 'PT_Elev', size = 0.1) + 
  tm_scale_intervals(breaks = 5,  
                     values = "brewer.yl_or_rd", 
                     style = "quantile") 

```
```{r}
## here we are using a simple dataset of the world 
# tmap_mode("plot")
data("World")
tm_shape(World) +
    tm_polygons("gdp_cap_est", style='quantile', legend.title = "GDP Per Capita Estimate")
```
```{r}
## the view mode creates an interactive map
tmap_mode("view")
tm_shape(World) +
    tm_polygons("gdp_cap_est", style='quantile', legend.title = "GDP Per Capita Estimate")
```

## Questions:

# 1.Discuss the advantages and challenges associated with an open data science approach. Provide an example based on this week’s reading. (1-2 paragraphs)

Open data science emphasizes transparency by enabling anyone to access, analyze, and reuse scientific observations, data, and results. This openness allows scientific claims to be validated by allowing other researchers to replicate and validate findings, which is essential for building trust in scientific conclusions. In addition, open data allows for the integration of datasets from multiple sources, resulting in new insights and discoveries that may not be possible with isolated datasets. For example, by integrating spatial data from the social media shape file with other environmental or population datasets, researchers can uncover patterns in urban development or environmental impacts that remain hidden without cross-referencing data.

However, open data science also brings challenges, including maintaining data privacy and ensuring that shared data is well documented and interoperable. Sensitive information must be protected, especially when dealing with personal data or vulnerable people. To reap the benefits of open data, researchers must ensure that data is formatted and annotated in a way that others can understand and use, which requires additional time and effort. For example, the social media shapefile dataset must contain clear documentation of any transformations in its structure, data fields, and applications to ensure that it can be effectively integrated and interpreted by other researchers, supporting reproducibility and potential new discoveries.

# 2.Create a markdown document that showcases an analysis of this week’s data or any other dataset of your choice. Include descriptive text that explains your analysis, and incorporate figures and geovisualizations.Include 1 chart and 1 map. Structure and explain your analysis with text, headings, highlights, images and other markdown basics.

## Analysis of Boulder Social Media Data

This document showcases an analysis of Boulder social media data, focusing on visualizing elevation distribution and exploring the relationship between elevation and proximity to bike trails.

### Importing and Inspecting Data

```{r}
# Load required packages
library(ggplot2)
library(tmap)
library(sf)
library(dplyr)
library(viridis)
library(tmaptools)
# Read the data
boulder <- st_read("/Users/zhongyinjiao/Desktop/EAS648/Lab01/BoulderSocialMedia.shp")
boulder
```
### Elevation Distribution Map

The following map uses the tmap package to visualize elevation across the Boulder area. Higher elevations are displayed in lighter shades, while lower elevations are in darker shades.

```{r}
library(ggplot2)
library(viridis)

ggplot() +
  geom_sf(data = boulder, aes(color = PT_Elev), fill = NA, alpha = 0.7) + 
  scale_colour_viridis_c(option = "magma", name = "Elevation") +
  labs(title = "Elevation Map of Boulder") +
  theme_minimal()
```
```{r}
library(ggplot2)
library(sf)
library(viridis)

# get the latitude and longtitude
boulder <- boulder %>% 
  mutate(x = st_coordinates(.)[,1],
         y = st_coordinates(.)[,2])

# plot the density
ggplot(data = boulder) +
  stat_density_2d(aes(x = x, y = y, fill = ..level..), geom = "polygon") +
  scale_fill_viridis_c(option = "magma") +
  labs(title = "Density Map of PT_Elev in Boulder") +
  theme_minimal()
```
```{r}
ggplot(data = boulder) +
  geom_point(aes(x = Bike_dis, y = PT_Elev, color = PT_Elev)) +
  scale_color_viridis_c(option = "inferno") +
  labs(title = "Bike Distance vs PT_Elev", x = "Bike Distance", y = "PT_Elev") +
  theme_minimal()
```

### Relationship Between Elevation and Bike Trail Distance

Here, I use ggplot2 to create a scatter plot to explore the relationship between elevation (PT_Elev) and the distance to the nearest bike trail (Bike_dis).
```{r}
# Plot elevation vs bike distance
ggplot(data = boulder) +
  geom_point(aes(x = Bike_dis, y = PT_Elev, color = PT_Elev), alpha = 0.6) +
  scale_color_viridis_c(option = "magma") +
  labs(title = "Bike Distance vs Elevation",
       x = "Distance to Bike Trails (Bike_dis)",
       y = "Elevation (PT_Elev)") +
  theme_minimal()

```

```{r}
# using ggplot2 to plot histogram of PT_Elev
ggplot(boulder, aes(x = PT_Elev)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  labs(
    title = "Elevation Distribution in Boulder",
    x = "Elevation (PT_Elev)",
    y = "Frequency"
  ) +
  theme_minimal()
```


### Analysis Summary

Elevation Distribution: The map highlights the variation in elevation across the Boulder area, with higher elevations concentrated in southwest regions.The scatter plot suggests that we can also see the elevation distribution from the histogram.
